---
title: "week 03 coding challenge"
author: "Talin Gharibian"
date: "02/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```


```{r load libraries}
library(tidyverse)
library(tidymodels)
```


```{r load data}
data.01 <- read_csv("Data/LHS_exercise_02042022.csv")

```

We want to explore whether higher risk of depression increases alcohol consumtion. 

Based on the Latino Health Survey codebook, the four items of depression (PTSD-4; page 6) need to be reversed.

Your task is to manipulate your data using the `tidyverse`'s `mutate`, `case_when` functions and run a series of logistic regression analyses. Use Google search as needed.

Challenge 1: mutate new variables: 
- `ptsd_mean` (average of four ptsd items)
- `ptsd_mean_rev` (reverse of the `ptsd_mean` variable; hint: 6- value)
- `ptsd_mean_rev_std` (standardized value of `ptsd_mean`)
- `audit_c` (participants who responded 1 or 2 ~ 1; else ~0)

For your reference: (pg 8 of the codebook)
How often did you have a drink containing alcohol in the past year? 
1. Four or more times a week <-- coded as '1' 
2. Two to three times per week <-- coded as '1'
3. Two to four times a month <-- coded as '0'
4. Monthly or less <- coded as '0'
5. Never <- coded as '0'

```{r data wrangling}
data.02 <- data.01%>%
  mutate(ptsd_mean = (ptsd1_1 + ptsd2_1 + ptsd3_1 + ptsd4_1)/4,
         ptsd_mean_rev = 6 - ptsd_mean,
         ptsd_mean_rev_std = scale(ptsd_mean_rev),
         audit_c = case_when(audit1 < 3 ~ 1,
                             audit1 > 2 ~ 0))
```



#Now we will be running several logistic regression and compare the results:

Run logistic regression using `ptsd_mean_rev` adjusting for age, sex, and education; use `tidy` function for output; exponent your results and interpret your result using odds ratio
```{r logistic regression 1}
fit.1 <- glm(audit_c ~ ptsd_mean_rev + age + factor(education) + factor(sex),
             family = binomial(link = "logit"),
             data = data.02)
summary(fit.1)
tidy(fit.1, conf.int = T, exponentiate =T)
```

Interpretation: 
[After controlling for education, age and sex, we find that depression risk (ptsd_mean_rev) is positively associated with alcohol consumption (audit_c), with an odds ratio of 1.36 and p<0.001.]

Use `ggfortify`'s `autoplot` function to do model diagnostics
```{r}
#install.packages('ggfortify')
library('ggfortify')
autoplot(fit.1)
```

Explain diagnostic plots:
[ The residuals deviate from the typical regression assumptions.]



Run logistic regression using `ptsd_mean_rev_std` adjusting for age, sex, and education; use `tidy` function for output; use `divide-by-four` approach and interpret your result using probability scale. 
```{r}
fit.2 <- glm(audit_c ~ ptsd_mean_rev_std + age + factor(education) + factor(sex),
             family = binomial(link = "logit"),
             data = data.02)
summary(fit.2)
tidy(fit.2, conf.int = T, exponentiate =T)
```
```{r}
#without exponentiate option:
data.03 <- tidy(fit.2, conf.int = T)
data.03$divdideby4 <- (data.03$estimate/4)*100
data.03
```

Interpretation
[ Using the standardized variable, we see that the ptsd_mean_rev_std and audit_c score are positively associated with an odds ratio of 1.39, p<0.001. Using the 'divide by 4 rule,' we find that increasing the ptsd_mean_rev_std score by one unit will increase the chance of the audit_c=1 score about 8.4%.]



#Next, visualization
Use `ggplot` to duplicate the figure
Using the `ptsd_mean_rev` variable you created before, you would need to categorize the mean value into four categories: 
if ptsd_mean_rev is :
1-1.9 ~ 1 ; 
2-2.9 ~2; 
3-3.9 ~ 3; 
4+ ~ 4. 

Refer to Healy's book or Google search as a guide. 

```{r visualization}
data.04 <- data.02%>%
  mutate(ptsd_mean_rev_cat = case_when(ptsd_mean_rev >= 4 ~ 4,
                             ptsd_mean_rev >= 3 ~ 3,
                             ptsd_mean_rev >= 2 ~ 2,
                             ptsd_mean_rev >= 1 ~ 1))
```

